Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes

### 2024-12-15 paper.qmd

| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
|--------------------------------|---------------------|--------------------|
| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |

### Question 1

@@

```{r, echo=FALSE, include=FALSE}
#looking for outliers 
library(car)
outlier_points <- outlierTest(mlr.imp)
print(outlier_points)
outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
print(outlier_indices)  # View the indices to confirm
```

@@

| Statistic          | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| Min                | 1.0                                 | 0.0                                             | 2.8                            | 28972                        | 0                        | 64.9                    |
| Median             | 50.3                                | 6.5                                             | 12.6                           | 58623                        | 1173                     | 75.1                    |
| Mean               | 49.3                                | 8.2                                             | 14.1                           | 61940                        | 1521                     | 75.2                    |
| Max                | 93.8                                | 62.6                                            | 47.1                           | 167605                       | 9788                     | 98.9                    |
| Standard Deviation | 16.5                                | 6.9                                             | 6.6                            | 16877.8                      | 1274.7                   | 3.1                     |

>>>>

| Variable | **Inadequate School Funding** | **Adequate School Funding** |
|----|----|----|
| Residential Segregation Index: Mean (SD) | 46 (17) | 56 (14) |
| Region-West: Count (%) | 114 (7.3%) | 89 (10.8%) |
| Region-Northeast: Count (%) | 3 (0.2%) | 200 (24.3%) |
| Region-Midwest: Count (%) | 267 (17.1%) | 285 (34.7%) |
| Region-South: Count (%) | 1173 (75.3%) | 248 (30.2%) |
| Median Property Tax, USD: Mean (SD) | 1119 (763) | 2282 (1648) |
| Median Household Income, USD: Mean (SD) | 56513 (13200) | 72218 (18255) |

=======
| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
|--------------------------------|---------------------|--------------------|
| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |

```{r}
# Plot for Q2 EDA
ggpairs(
  completed_data,
  columns = c("Residential.Segregation.Index", "Median.Household.Income", "Median.Prop.Tax"),
  aes(color = factor(School.Funding.Cat,
                     labels = c("Inadequate", "Adequate")))
) +
  labs(
    title = "Pairwise Plots of Predictors Colored by School Funding Adequacy",
    color = "School Funding Adequacy"
  ) +
  theme(
    plot.title = element_text(size = 15, hjust = 0.5),
    strip.text.x = element_text(size = 5),
    strip.text.y = element_text(size = 3)
  )

```
### Question 1

@@

```{r, echo=FALSE, include=FALSE}
#looking for outliers 
outlier_points <- outlierTest(mlr.imp)
print(outlier_points)
outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
print(outlier_indices)  # View the indices to confirm
```

@@

| Statistic | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
|----|----|----|----|----|----|----|
| Min | 1.0 | 0.0 | 2.8 | 28972 | 0 | 64.9 |
| Median | 50.3 | 6.5 | 12.6 | 58623 | 1173 | 75.1 |
| Mean | 49.3 | 8.2 | 14.1 | 61940 | 1521 | 75.2 |
| Max | 93.8 | 62.6 | 47.1 | 167605 | 9788 | 98.9 |
| Standard Deviation | 16.5 | 6.9 | 6.6 | 16877.8 | 1274.7 | 3.1 |
